A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-11-19 paper/paper.tex
paper: tweak table numbering
About parametricity

% The eqntable environment
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\NewEnviron{eqntable}[1]{
  \scriptsize
  \begin{tabularx}{0.9\linewidth}{
    @{}
    >{$}l<{$}
    >{$}c<{$}
    >{$}c<{$}
    >{$}Y<{$}
    @{}}
    \toprule
    \BODY
    \bottomrule
  \end{tabularx}
  \captionof{table}{#1}
}

@@

In there, the process of construction of the type of $X_1$ from that of $X_0$, and from the type of $X_2$ to that of $X_1$, and so on, is similar to iteratively applying a binary parametricity translation. The parametricity which we consider interprets a type $A$ by a family $A_\kstar$ over $A \times A$, what can be seen as a graph whose vertices are in $A$. To each type constructor is thus associated the construction of a graph. To start with, the type of types $\U$ is interpreted as the family of type of families ${\U}_\kstar$, which takes $A_L$ and $A_R$ in $\U$ and returns the type $A_L \times A_R \rightarrow \U$ of families over $A_L$ and $A_R$. Also, for $A$ interpreted by $A_\kstar$ and $B$ interpreted by $B_\kstar$, a dependent function type $\Pi a: A.\, B$ is interpreted as the graph $(\Pi a: A.\, B)_\kstar$ that takes two functions $f_L$ and $f_R$ of type $\Pi a: A.\, B$, and expresses that these functions map related arguments in $A$ to related arguments in $B$:
\begin{align*}
  (\Pi a: A.\, B)_\kstar(f_L,f_R) \; \defeq \; \Pi (a_L,a_R): (A \times A).\, (A_\kstar(a_L,a_R)\,\rightarrow B_\kstar(f_L(a_L),f_R(a_R)))
\end{align*}

@@

In particular, for $X: \U$, applying the parametricity construction is about canonically associating to $X$ an inhabitant $X_\kstar$ of ${\U}_\kstar(X,X)$ i.e. of $X \times X \rightarrow \U$. In turn, applying the parametricity translation to $X_\kstar: X \times X \rightarrow \U$ is about canonically associating to $X_\kstar$ an inhabitant $X_{\kstar\kstar}$ of $(X \times X \rightarrow \U)_\kstar(X_\kstar,X_\kstar)$ i.e. of:
\begin{align*}
  \Pi ((x_{LL},x_{lR}),(x_{RL},x_{RR})): ((X \times X) \times (X \times X)). \\
  (X_\kstar(x_{LL},x_{LR}) \times X_\kstar(x_{RL},x_{RR})
  \rightarrow X_\kstar(x_{LL}, x_{RL}) \times X_\kstar(x_{LR},x_{RR})  \rightarrow \U)
\end{align*}
which hints us at how the sequence $X_0$, $X_1$, $X_2$ can be seen as a sequence of inhabitants of the iteration of binary parametricity applied to an initial $X: \U$:
\begin{equation*}
  \begin{array}{llllr}
    X_0 & \defeq & X                & : & \U                                             \\
    X_1 & \defeq & X_\kstar         & : & {\U}_\kstar(X,X)                               \\
    X_2 & \defeq & X_{\kstar\kstar} & : & ({\U}_\kstar(X,X))_{\kstar}(X_\kstar,X_\kstar) \\
    \ldots
  \end{array}
\end{equation*}

@@

% Insert fulltab:faces and fulltab:coh, after overriding macros to make
% everything but the universe letter explicit.
\renewcommandx{\framep}[5][1,2,3,4,5]{\prim{frame}[][#2][#3][#4][#5]}
\renewcommandx{\layer}[5][1,2,3,4,5]{\prim{layer}[][#2][#3][#4][#5]}
\renewcommandx{\painting}[5][1,2,3,4,5]{\prim{painting}[][#2][#3][#4][#5]}
\renewcommandx{\restrf}[7][1,2,3,4,5,6,7]{\restr{frame}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\restrl}[7][1,2,3,4,5,6,7]{\restr{layer}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\restrc}[7][1,2,3,4,5,6,7]{\restr{painting}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\cohf}[9][1,2,3,4,5,6,7,8,9]{\coh{frame}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohl}[9][1,2,3,4,5,6,7,8,9]{\coh{layer}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohc}[9][1,2,3,4,5,6,7,8,9]{\coh{painting}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\coht}[9][1,2,3,4,5,6,7,8,9]{\cohtwo{frame}[][#2][#3][#4][#5][#6][#7,#8][#9]}
\renewcommand*{\lab}{fulltab:faces}
\input{tab-faces.tex}
\renewcommand*{\lab}{fulltab:coh}
\input{tab-coh.tex}

>>>>

% The eqntable environment
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\NewEnviron{eqntable}[1]{
  \scriptsize
  \begin{tabularx}{0.93\linewidth}{
    @{}
    >{$}l<{$}
    >{$}c<{$}
    >{$}c<{$}
    >{$}Y<{$}
    @{}}
    \toprule
    \BODY
    \bottomrule
  \end{tabularx}
  \captionof{table}{#1}
}

@@

In there, the process of construction of the type of $X_1$ from that of $X_0$, and from the type of $X_2$ to that of $X_1$, and so on, is similar to iteratively applying a binary parametricity translation. The binary parametricity which we consider is composed with the diagonal on types and interprets a type $A$ by a family $A_\kstar$ over $A \times A$, what can be seen as a graph whose vertices are in $A$. To each type constructor is thus associated the construction of a graph. To start with, the type of types $\U$ is interpreted as the family of type of families ${\U}_\kstar$, which takes $A_L$ and $A_R$ in $\U$ and returns the type $A_L \times A_R \rightarrow \U$ of families over $A_L$ and $A_R$. Also, for $A$ interpreted by $A_\kstar$ and $B$ interpreted by $B_\kstar$, a dependent function type $\Pi a: A.\, B$ is interpreted as the graph $(\Pi a: A.\, B)_\kstar$ that takes two functions $f_L$ and $f_R$ of type $\Pi a: A.\, B$, and expresses that these functions map related arguments in $A$ to related arguments in $B$:
\begin{align*}
  (\Pi a: A.\, B)_\kstar(f_L,f_R) \; \defeq \; \Pi (a_L,a_R): (A \times A).\, (A_\kstar(a_L,a_R)\,\rightarrow B_\kstar(f_L(a_L),f_R(a_R)))
\end{align*}

@@

In particular, for $X: \U$, applying our parametricity translation is about associating to $X$ an inhabitant $X_\kstar$ of ${\U}_\kstar(X,X)$ i.e. of $X \times X \rightarrow \U$. In turn, applying the translation again to $X_\kstar: X \times X \rightarrow \U$ is about associating to $X_\kstar$ an inhabitant $X_{\kstar\kstar}$ of $(X \times X \rightarrow \U)_\kstar(X_\kstar,X_\kstar)$ i.e. of:
\begin{align*}
  \Pi ((x_{LL},x_{lR}),(x_{RL},x_{RR})): ((X \times X) \times (X \times X)). \\
  (X_\kstar(x_{LL},x_{LR}) \times X_\kstar(x_{RL},x_{RR})
  \rightarrow X_\kstar(x_{LL}, x_{RL}) \times X_\kstar(x_{LR},x_{RR})  \rightarrow \U)
\end{align*}
which hints us at how the sequence $X_0$, $X_1$, $X_2$ can be seen as a sequence of inhabitants of the iteration of the composition of binary parametricity with the diagonal on types and type families, applied to an initial $X: \U$:
\begin{equation*}
  \begin{array}{llllr}
    X_0 & \defeq & X                & : & \U                                             \\
    X_1 & \defeq & X_\kstar         & : & {\U}_\kstar(X,X)                               \\
    X_2 & \defeq & X_{\kstar\kstar} & : & ({\U}_\kstar(X,X))_{\kstar}(X_\kstar,X_\kstar) \\
    \ldots
  \end{array}
\end{equation*}

@@

% Insert fulltab:faces and fulltab:coh, after overriding macros to make
% everything but the universe letter explicit.
\renewcommandx{\framep}[5][1,2,3,4,5]{\prim{frame}[][#2][#3][#4][#5]}
\renewcommandx{\layer}[5][1,2,3,4,5]{\prim{layer}[][#2][#3][#4][#5]}
\renewcommandx{\painting}[5][1,2,3,4,5]{\prim{painting}[][#2][#3][#4][#5]}
\renewcommandx{\restrf}[7][1,2,3,4,5,6,7]{\restr{frame}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\restrl}[7][1,2,3,4,5,6,7]{\restr{layer}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\restrc}[7][1,2,3,4,5,6,7]{\restr{painting}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\cohf}[9][1,2,3,4,5,6,7,8,9]{\coh{frame}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohl}[9][1,2,3,4,5,6,7,8,9]{\coh{layer}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohc}[9][1,2,3,4,5,6,7,8,9]{\coh{painting}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\coht}[9][1,2,3,4,5,6,7,8,9]{\cohtwo{frame}[][#2][#3][#4][#5][#6][#7,#8][#9]}
\renewcommand*{\lab}{fulltab:faces}
\renewcommand{\thetable}{4'}
\input{tab-faces.tex}
\renewcommand*{\lab}{fulltab:coh}
\renewcommand{\thetable}{5'}
\input{tab-coh.tex}
